Rough Analysis

### 2023-10-06 Lahman Analysis/Rough Analysis/RoughAnalysis.qmd
Updated Model

install.packages("Lahman")
library(Lahman)
```

@@

```{r}
#There are more datasets in Lahman than we care about so we'll select "Batting"

Batting <- Lahman::Batting
```

@@

```{r}
Modern.Batting <- Batting %>% 
  filter(yearID > 1945)
```

@@

```{r}
#Of course we need the other two main libraries
library(tidyverse)
library(rethinking)
```

@@

@@

>>>>

install.packages("Lahman")
```

@@

```{r}
#There are more datasets in Lahman than we care about so we'll select "Batting"
library(tidyverse)
library(Lahman)

Teams <- Lahman::Teams
```

@@

```{r}
Modern.Teams <- Teams %>% 
  filter(yearID > 1945)
```

@@

```{r}
#| warning: false
#Of course we need the other main libraries
library(rethinking)
```

# What are we modeling?

For this project, I want to model how many homeruns a team hits in a season. We know that this should probably be normally distributed and we can derive our expected mean $\mu$ with a little math.

In the more modern game, having a 25HR would probably be considered about average. With 9 guys on the team with some above and some below then we could expect about $9\cdot25=225$. So we would expect teams to hit about 225 homeruns in a year. However, our sample is of teams from 1945 to 2022 which is decidedly not comprised of modern teams. Furthermore, in the "good-ol-days" batting average was stressed much more than power; hence players did not hit as many home runs. Therefore, lets guess that each player averaged about 15HR a season. This would mean we could expect each team to hit $9\cdot15=135$ homeruns each year.

Now with our general guess of the mean we can use a prior. However, this problem is shaping up as one that might lend itself not to a Gaussian prior, but a Poisson one.

Furthermore, we want to model this, not just match distributions. So, I would argue (since we're keeping this simple) that the number of team AB's has a lot to do with how many home runs the team hits. More AB's implies more chances to hit HR's and suggests the team is just better at not getting out overall. So, let's use AB's as our predictor.

@@

```{r}
#Calculate mean AB's now so it's easy to access later
mean_AB <- mean(Modern.Teams$AB)
```

```{r}
q1 <- quap(
  alist(
    HR ~ dnorm(mu, sigma),
    mu <- a + b*(AB - mean_AB),
    #Standard deviation of 20 seems fair to me for home runs in a year
    a ~ dnorm(135, 20),
    #Want to make a very non descriptive prior here since I don't know what the value of b could be
    b ~ dnorm(1, 1),
    sigma ~ dexp(1)
  ), data = Modern.Teams
)
```

```{r}
precis(q1, depth=2)
```

@@

```{r}
u1 <- ulam(
  alist(
    HR ~ dnorm(mu, sigma),
    mu <- a + b*(AB - mean_AB),
    #Standard deviation of 20 seems fair to me for home runs in a year
    a ~ dnorm(135, 20),
    #Want to make a very non descriptive prior here since I don't know what the value of b could be
    b ~ dnorm(1, 1),
    sigma ~ dexp(1)
  ), data = Modern.Teams, chains=1
)
```

```{r}
precis(u1)
```

# Comparison
